---
title2: "P8105_hw1_yc384"
author: "Ying Chen (UNI: yc384)"
date: "9/14/2019"
output: github_document
---

# **P8105 DSI Homework 1**

## Problem 0.1: Use R Markdown to write reproducible reports, GitHub for version control, and R Projects to organize your work.

* Created a public GitHub repo and a local R Project both named P8105_HW1_yc384
* Created a single .Rmd file: P8105_HW1_yc384.Rmd
* My link to homework1 repo is: https://github.com/YingCarolineChen/P8105_HW1_yc384.git


## Problem 0.2: Code styling: pay attention the the points listed below

* Meaningful variable / object names
* Readable code (one command per line; adequate whitespace and indentation; etc)
* Clearly-written text to explain code and results
* A lack of superfluous code (no unused variables are defined; no extra library calls; ect)


## Problem 1: Understand variable types in R: numeric, character, and factor variables:

### 1_1. First code chunk: This chunk Create a data frame comprised of:
* a random sample of size 8 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 8
* a factor vector of length 8, with 3 different factor “levels”

```{r HW1_1_df, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

## Create a dataframe 
hw1_df = tibble(
 ## A random sample of size 8 from a standard Normal distribution
 norm_samp = rnorm (8, mean = 0, sd = 1),
 ## Logic vector: if elements of the random sample are positive
 norm_samp_pos = norm_samp > 0,
 ## Character vector of length 8
 vec_char = c("a", "b", "c", "abc", "ad", "ac", "bc", " "),
 ## Factor vector of length 8 with 3 differesnt levels
 vec_factor = factor(c("low", "high", "medium", "high", "low", "medium", "high",  
                       "low")),
)

## Take the mean of each variable
## Examing the mean of varible norm_samp: 
mean(hw1_df$norm_samp)
## Examing the mean of varible norm_samp_pos: 
mean(hw1_df$norm_samp_pos)
## Examing the mean of varible var_char: 
mean(hw1_df$vec_char)
## Examing the mean of varible var_factor: 
mean(hw1_df$vec_factor)

```

* **From results above, we can see that the ransom normal sample and the logic vector had their mean calculated. But the character variable and the factor variable did not have the mean calculated.**


### 1_2.  Second code chunk applies as.numeric function to the other types of varaibles(only show code, not output)
        
```{r, eval=FALSE}
as.numeric(pull(hw1_df, norm_samp_pos))
as.numeric(pull(hw1_df, vec_char))
as.numeric(pull(hw1_df, vec_factor))

```
* When apply as.numeric function to the logical variables, it returns 0 and 1; 
* When apply as.numeric function to the character variables, it returns NA;
* When apply as.numeric function to the factor variables, it returns three levels in   
  numbers according to their alphabet order;

* Yes. This is helpful to explain why we could not take the mean of a character or a factor variables. They are not numeric variables. 

### 1_3. Third code chunk:  
* convert the logical vector to numeric, and multiply the random sample by the result
* convert the logical vector to a factor, and multiply the random sample by the result
* convert logical vector to factor & then convert result to numeric & multiply the random sample by the result

```{r, include = TRUE}
## Create a data frame that converts different types of variable
hw1_dt = tibble(
  norm_samp = rnorm (8, mean = 0, sd = 1),
  
  ## Convert logical vector to numeric
  logic_num = as.numeric(pull(hw1_df, norm_samp_pos)),
  ## multiply converted variable by the random sample
  logic_num_normsamp = logic_num * norm_samp,
  
  ## Covert logical vector to factor
  logic_factor = as.factor(pull(hw1_df, norm_samp_pos)),
  ## multiply converted variable by the random sample
  logic_factor_normsamp = logic_factor * norm_samp,
  
  ## convert logical vector to factor & then convert result to numeric
  logic_factor_num = as.numeric(as.factor(pull(hw1_df, norm_samp_pos))),
  ## multiply converted variable by the random sample
  logic_factor_num_normsamp = logic_factor_num * norm_samp
)
 
```

## Problem 2: Using ggplot to plot different types of variables

### 2_1 Create data frame: HW1_2_df
```{r HW1_2_df, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

## set seed to ensure reproducibility
set.seed(1)

## Create a dataframe 
HW1_2_df = tibble(
## A random sample of size 500
 x = rnorm (500),
 y = rnorm (500),

 # Logic vector: if x + y > 1
 Logical_var1 = x + y > 1,
 # coercing the above logical vector to numeric
 numeric_var1 = as.numeric(Logical_var1),
 # coercing the above logical vector to factor
 factor_var1 = as.factor(Logical_var1),
)

```

### 2_2 Description of Dataset HW1_2_df: 
* The dataset: hw2_df contains two random samples of size `r nrow(HW1_2_df)` and `r ncol(HW1_2_df) ` variables. 
* The mean of x is `r mean(pull(HW1_2_df,x))`, the median of x is `r median(pull(HW1_2_df, x) )` and standard deviation of x is `r sd(pull(HW1_2_df, x))`
* the proportion of cases for which x + y > 1 is: `r sum(pull(HW1_2_df, numeric_var1))/500*100`%.

### 2_3 Make a scatterplot of y vs x
```{r HW1_2_plots, include=TRUE}

## Scatter plot for logical variable
ggplot(HW1_2_df, aes(x = x, y = y, color=Logical_var1)) + geom_point()
## Export the first scatterplot to project directory using ggsave
ggsave("plot1.jpeg", plot = last_plot())

## Scatter plot for numerical variable, color by factor_var1
ggplot(HW1_2_df, aes(x = x, y = y, color=numeric_var1)) + geom_point()

## Scatter plot for factor variable, color by numerical varaible
ggplot(HW1_2_df, aes(x = x, y = y, color=factor_var1)) + geom_point()

```
